# Getting Started with R{#getting-started-with-r}

Over the last few years, two approaches have emerged to deal with issues related to the quantitative analysis of text: R and Python. While Python is a general-purpose language (useful for everything from websites to games to databases), R was always designed with statistics in mind. As a result, R is better suited to deep statistical analysis and a wide range of data visualisation. As a result, while Python is a good choice when dealing with issues related to machine learning and large-scale applications, R is better suited to statistical learning and data exploration. Since we are mainly interested in the latter here, we built this book around R (and all the packages it offers). If you are more interested in large-scale analysis using Python, be sure to take a look at its [spaCy](https://spacy.io) and [NLTK](https://www.nltk.org) libraries.

## What is R?

R is an open-source program that allows you to perform a wide variety of statistical tasks. At its core, it is a modification of the S and Scheme programming languages, making it not only flexible but also fast. R is available for Windows, Linux and OS X, and is updated regularly. In its basic form, R uses a simple command line interface. To make it more friendly, you can use one of the integrated development environments (IDEs) such as [RStudio](https://www.rstudio.com/), [Jupyter](https://jupyter.org/) and [R Commander](https://socialsciences.mcmaster.ca/jfox/Misc/Rcmdr/). These environments not only look better but also offer some additional practical features. In this book, we will use RStudio as it is (in our opinion) the best-looking and working IDE available.

## R on Windows

To install R on Windows, go to https://cran.r-project.org/bin/windows/base/, download the file, double-click it and run it. During installation, it is best to leave default options (such as the installation folder) unchanged. This makes it easier for other programs to know where to find R. Once installed, you will find two shortcuts for R on your desktop. These refer to the two versions of R that come with the installation - the 32-bit version and the 64-bit version. Which version you need will depend on your version of Windows. To find out which version of Windows you have, go to **This PC** (or **My Computer**), right-click and select **Properties**. Here you should find the version of Windows installed on your PC. If you have a 64-bit version of Windows, you can use either version. However, it is best to use the 64-bit version as it makes better use of your computer's memory and runs more smoothly. If you have the 32-bit version of Windows, you must use the 32-bit version of R.

To install RStudio, go to https://www.rstudio.com/products/rstudio/download/ and download the free version of RStudio at the bottom of the page. Make sure you select **Installers for Supported Platforms** and select the option for Windows. Once downloaded, install the program leaving all settings unchanged. If all goes well, RStudio will have found your installation of R and placed a shortcut on your desktop. Whether you have a 32-bit or 64-bit version of Windows or R does not matter to RStudio. What does matter are the slashes. R uses slashes (`/`) instead of the backslashes (`\`) that Windows uses. So whenever you refer to a folder or file in R, make sure you reverse the slashes. For example, you should refer to a file that has the Windows address **C:\Users\\Desktop\\data.csv** as **C:/Users/Desktop/data.csv**.

## R on Linux

How you install R on Linux depends on which flavour of Linux you have. In most cases, R is already included in your Linux distribution. You can check this by opening a terminal and typing `R`. If it is installed, R will start in the terminal. If R is not part of your system, run the following in a terminal

1. sudo apt update
2. sudo apt install r-base r-base-dev -y
3. Now run "R" to see if the installation worked

To install RStudio, go to https://www.rstudio.com/products/rstudio/download/. At the bottom of the page, select the installer for your operating system. Then install the file either using an installation manager or via the terminal. After running the launcher, you will find RStudio in the Dash.

## R on MacOS

For OS X, you must have OS X 10.6 (Snow Leopard) or later. Otherwise, you can still install R, but you will not be able to use a certain number of packages (like some we use here). To check this, click on the Apple icon at the top left of your screen. Then click on the **About This Mac** option. A window should appear telling you which version of OS X (or MacOS) you have.

To install R, go to https://cran.r-project.org/index.html and click on **Download R for (Mac) OS X**. Once there, download and install the appropriate .pkg file for your version of OS X. You will also need to download the **Clang 6.x Compiler** and the **GNU Fortran Compiler** from https://cran.r-project.org/bin/macosx/tools/. Install both, leaving the selected options as they are. Once installed, check that R works by running it.

To install RStudio, go to https://www.rstudio.com/products/rstudio/download/ and download the OSX version at the bottom of the page. Once downloaded and installed, you should be able to find RStudio among your other applications.

## R in the Cloud

As well as installing R on your system, you can also use the cloud version. RStudio hosts this version at https://posit.cloud/. To use it, go to the Sign-Up button at the top right of the screen. Then select the **Cloud Free** option and select Sign-Up again. Finish the process by either filling in your details or connecting to your Google or GitHub account. Once done, log in and you will be taken to your workspace. To get started, you will need to create a new project. To do this, click on the **New Project** button, which will take you to an instance of RStudio. From here on, the program works in exactly the same way as the desktop version. Note that anything you do - or packages you install - in the project will remain in the project. So if you want to create a new project, you will have to reinstall it. Also, note that Posit Cloud is highly dependent on both the number of users on the server and your internet connection. Therefore, some actions (such as installing packages) may take longer to complete.


# Installing Packages{#installing-packages}

R on its own is a pretty bare-bones experience. What makes it work are the many packages that exist for it. These packages come in two types: those with an official release and those in development.

## Installing from CRAN

For an official release, the package must be part of CRAN: the *Comprehensive R Archive Network*.CRAN is a website that collects and hosts all the material R needs, such as the various distributions, packages and more. The main advantage of being official is that it means the package has been through a review process. This ensures that the package is free of major bugs, has README and NEWS files, and has a unique version number. In addition, packages that have been officially released often have accompanying articles in [*The R Journal*](https://www.https://journal.r-project.org/) or [*The Journal of Statistical Software*](https://www.jstatsoft.org/index) that provide detailed explanations of the code, examples, etc.

To install official packages, we can use the `install.packages' command or the **Packages** tab in RStudio. If we run `install.packages("package", dependencies=TRUE)`, we will also install any other packages a package depends on. Finally, to update, we can either go to the **Packages** tab in RStudio and click the **Update** button. We can also type `update.packages()` in the **Console ** for the same result. Try to update your package often to avoid unnecessary bugs and problems.

## Installing from GitHub

Some packages that have not yet had an official release are in development on [**GitHub**](https://github.com/). As a result, these packages change very often and are more unstable than their official counterparts. We can install packages from Github using the `devtools` package. To install this, type:

```{r install-devtools, eval=FALSE}
install.packages("devtools", dependencies=TRUE)
```

Here, `dependencies=TRUE` ensures that if we need other packages to make `devtools` work, R will install these as well. Depending on your operating system, you might have to install some other software for `devtools` to work.

On Windows, `devtools` requires the **RTools** software. To install this, go to https://cran.r-project.org/bin/windows/Rtools/, download the latest **recommended** version (in green), and install it. Then re-open R again and install `devtools` as shown above.

On Linux, the way you install `devtools` depends on the flavour of Linux you have. In most cases, installing `devtools' from the RStudio console will work fine. If not, the problem is most likely a missing package in your Linux distribution. To fix this, close R, open a terminal and type:

1. `sudo apt-get update`
2. `sudo apt-get upgrade`
3. `sudo apt install build-essential libcurl4-gnutls-dev libxml2-dev libssl-dev`
4. Close the terminal, open R, and install `devtools` as shown above.

On OSX (or macOS), `devtools` requires the *XCode* software. To install this, follow these steps:

1. Launch the terminal (which you can find in */Applications/Utilities/*), and type:
2. In the terminal, type: `xcode-select --install`
3. A software update window should pop up. Here, click **Install** and agree to the Terms of Service.
4. Go to R and install `devtools` as shown above.

## Packages for Quantitative Text Analysis in R

There are several packages that we can use for quantitative text analysis in R, such as `tm`, `tidytext`, `RTextTools`, `corpus` and `koRpus` [@Welbers2017a]. Many of these packages offer specialised features that can sometimes be very useful, but in this book, we will mainly rely on `quanteda` [@Benoit2018a], which is currently in its fourth version. The advantage of `quanteda` over other packages is that it integrates into a common framework many of the text analysis functions of R that were previously spread across many different packages[@Welbers2017a].In addition, many `quanteda' functions can be easily combined with functions in other packages, while the package as a whole has simple and logical commands and a well-maintained [website](www.quanteda.io).

The current version of `quanteda` at the time of writing is 4.0. This version works best with R version 4.0.1 or higher. To check if your system has this, type `R.Version()` in your console. The result will be a list. Look for `$version.string` to see what version number your version of R is. If you do not have the latest version, see the steps above to install the latest version.

To install `quanteda`, type:

```{r install-quanteda, eval=FALSE}
install.packages("quanteda", dependencies = TRUE)
```

Note that because we wrote `dependencies = TRUE`, this command also installed three other `quanteda` helper packages, which serve to extend the basic tools that are already inside `quanteda`. In the future, more of these helper packages can be expected to extend the main `quanteda` package even further. However, before these helper packages get an official release, we can already find them in development on GitHub.In this book, we will install two of them - `quanteda.classifiers`, which we will use for supervised learning methods, and `quanteda.dictionaries`, which we will use for dictionary analysis:

```{r install-quanteda-helpers, eval=FALSE}
library(devtools)
install_github("quanteda/quanteda.classifiers", dependencies = TRUE)
install_github("kbenoit/quanteda.dictionaries", dependencies = TRUE)
install_github("quanteda/quanteda.corpora", dependencies = TRUE)
```

In addition to `quanteda` we then use the following packages:

```{r install-packages, eval=FALSE}
install_github("mikegruz/kripp.boot", dependencies = TRUE)
install.packages("ca", dependencies = TRUE)
install.packages("combinat", dependencies = TRUE)
install.packages("DescTools", dependencies = TRUE)
install.packages("FactoMineR", dependencies = TRUE)
install.packages("factoextra", dependencies = TRUE)
install.packages("Factoshiny", dependencies = TRUE)
install.packages("Hmisc", dependencies = TRUE)
install.packages("httr", dependencies = TRUE)
install.packages("jsonlite", dependencies = TRUE)
install.packages("manifestoR", dependencies = TRUE)
install.packages("readr", dependencies = TRUE)
install.packages("readtext", dependencies = TRUE)
install.packages("reshape2", dependencies = TRUE)
install.packages("RTextTools", dependencies = TRUE)
install.packages("R.temis", dependencies = TRUE)
install.packages("rvest", dependencies = TRUE)
install.packages("seededlda", dependencies = TRUE)
install.packages("stm", dependencies = TRUE)
install.packages("tidyverse", dependencies = TRUE)
install.packages("topicmodels", dependencies = TRUE)
install.packages("magick", dependencies = TRUE)
install.packages("vader", dependencies = TRUE)
```

Some of these are specialised packages for text analysis, others for statistical estimation and visualisation. After installation, you will find these packages, as well as the `quanteda` and `devtools` packages, under the **Packages** tab in RStudio.

## Problems, Bugs and Errors

It is not uncommon to get errors when typing a command in R. Errors often occur when you misspell the code or use the wrong code for the job at hand. In these cases, R will print a message (in red) telling you why it cannot do what you want it to do. Sometimes this message is quite clear, like telling you to install an extra package. Other times it is more complicated and requires some extra work. In these cases, there are four questions you should ask yourself:

1. Have I downloaded all the packages I need?
2. Are all my packages up to date?
3. Did I spell the commands correctly?
4. Is the data in the correct form or format?

If none of these provide a solution, you can always look online to see if others have encountered the same problem. Often, copying and pasting your error into a search engine will give you other examples, and usually a solution. A well-known place for solutions is [Stack Overflow](https://stackoverflow.com/), where you can share your problem with others and see if someone can offer a solution. However, make sure you read through the problems that have already been posted to make sure you do not post the same problem twice. We have also compiled a list of problems encountered by people who have used this book, along with their solutions, in this spreadsheet.



